The effect of symbols in government communications on citizens' trust, and its interaction with substantive information

### 2018-06-16 paper2_information and symbols_01.Rmd
Add files via upload

---
title: "The effect of symbols in government communications on citizens' trust, and its interaction with substantive information"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
link-citations: yes
bibliography: phd_paper_2.bib
urlcolor: blue
---

@@

Last edited at `r Sys.Date()`)[^footnote_scripts].

@@

Elements in communications can also function as factors that affect people's motivation and/or ability to think. For example, Background music in commercials can distract people and prevent them from scrutinizing their content [@park_1986]; The use of humor in television shows stimulates positive mood and reduces people's ability and motivation to scrutinize their substantive arguments [@nabi_2007; @young_2008]. Hence, elements can have more than one function, and can shape attitudes through several mechanisms. They can serve both as variables that affect the amount of thinking and as potential peripheral cues or arguments. Continuing with the above-mentioned examples, background music and humor can serve as simple cues that persudate people directly through the peripheral route, yet at the same time they can also as factors that reduce elaboration. Accordingly, they can affect people both directly (as peripheral cues) and indirectly (by decreasing elaboration and thus attenuating the effect of arguments and enhancing the effect of peripheral cues).[^footnote_multifunction_factors]

[^footnote_multifunction_factors] : @petty_1986 have suggested that we are more likely to find complex relations of multiple mechanisms in those situations where people are located in the middle of the elaboration continuum. When people are neither extremely motivated or unmotivated to think, they may rely on different elements in the communications to decide if the message is worth or needs considering.

@@

Analytically, the expectation that symbolic elements decrease message elaboration also entails that symbols may decrease citizens' trust in communications about highly positive performance and persuasive policy plans. Nevertheless, one can make the argument that neutralizing the effect of "positive" information with symbols may not actually yield significant differences, since citizens are more likely to "punish" bureaucracies for publishing negative information than to "praise" them for publishing positive information [@james_2011; @nielsen_2016; @james_2017b]. Congruently, the effect of positive information (versus a medium level) is likely to be relatively small, and hence neutralizing this effect may not make much difference. Given the focus of this study on the interaction between symbols and ill-conceived policies, the research design does not directly addresses this latter interaction.

@@

In addition to the abovementioned hypotheses, the experiment is also designed to test an additional set of hypotheses regarding the interaction between symbols and the perceived personal relevance of communications. The selection of policy plans and the construction of the population sample were designed to account for variation in that variable. In this paper, I focus on those methodological and empirical parts of the study that are consequential for the examination of the interaction bewteen symbols and information.[^footnote_relevance_paper]

[^footnote_relevance_paper]: The methodological aspects which are relevant for examining the additional hypotheses (i.e. the interaction between symbols with personal relevance) are described in a seperate paper. An English translation of the full survey is available from the supplementary appendix.

@@

[^footnote_prior_studies]: In two prior online survey tests, which I conducted before the experiment, I collected further information about the affect and symbolic associations of these symbols. These results, which are reported in the supplementary appendix [##TBA##], substantiate my empirical assumption that these elements evoke positive emotions Israeli citizens, and that citizens automatically associate them with EPM and its policies. These results also indicate that the images of the celebrity comedians from EPM campaigns are perceived as humoristic, and accordingly, they are expected to position people in a positive mood.

@@

[^footnote_relevance_manipulation]: This section also included a manipulation of participants' perceived personal relevance. Participants in the treatment group were asked a series of questions about environmental governent policy, whereas participants in the control group were asked instead equivalent questions about their field of occupation.

@@

[^footnote_symbols_manipulation_checks]: In two prior surveys, I empirically confirmed that the fake symbols are unfamiliar, do not have strong positive or negative affect, and are not strongly associated with environment policy and EPM. The results of these examinations are reported in the supplementary appendix (##TBA##).

@@

[^footnote_combined_analyses]: In the supplementary appendix I further replicate the analyses below, while separating between the two policies.

@@

My main independent variables are the above mentioned manipulations for information and symbols, and the interaction bewteen them. My outcome variable is citizens' trust in policy (or the perceived trustworthiness of the policy). This variable is measured in the survey for each policy plan using a composite index of six likert scale items (Cronbach’s alpha = `r alpha_waste_trust`), which were adopted and slightly modified from validated trust scales used by previous public administration studies on trust in government [@grimmelikhuijsen_2013; @grimmelikhuijsen_2014; @karens_2016; @porumbescu_2016; @grimmelikhuijsen_2017]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7(strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan. The mean of this trust index is `r SVIVA2_01_comb%>%summarise(mean(trust))%>%max()%>%round(2)`, the standard variation is `r SVIVA2_01_comb%>%summarise(sd(trust))%>%max()%>%round(2)` and the median is `r SVIVA2_01_comb%>%summarise(median(trust))%>%max()%>%round(2)`. Higher levels of trust are associated with high trust vis a vis government and left-wing ideology, as expected. In addition to measuring pariticipants trust in each policy, per each subpect, I also calculated the delta between her trust in the strong and weak policy plans which were presented to her (*Mean* = `r SVIVA2_01%>%summarise(mean(WITHIN_DELTA))%>%max()%>%round(2)`, *SD* = `r SVIVA2_01%>%summarise(sd(WITHIN_DELTA))%>%max()%>%round(2)`, *Median* = `r SVIVA2_01%>%summarise(median(WITHIN_DELTA))%>%max()%>%round(2)`). I used this Trust in policy delta to analyze the interaction between symbols and information within subjects.

@@

In addition to the above independent variables, the survey also included control items regarding respondents’ age, gender, education, income, home ownership, age of their children, political ideology, trust in governemnt, interest in environmental issues and residence in the Haifa-Bay area. Summary statistics for all our variables are available in the supplementary appendix.

@@

I now turn to analyzing the experimental results. I begin by examining my first and second hypotheses, that regard the effect of symbols on trust and its interaction with information. Thereafter, I will exmaine hypotheses 3 and 4, that regard the mediating role on elaboration. For each section, I conducted both *between-subjects* and *within-subjects* analyses.

@@

**Table 1: Regression analyses**

@@

stargazer(mod_1.1,mod_1.2,mod_1.3,
          type="html",style = "apsr",
          dep.var.labels=c("Trust in policy (combined)"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Weak policy (0=strong policy)",
                             "Weak policy x Real symbols",
                             "Weak policy x Fake symbols"),
          initial.zero = FALSE,
          notes = "*Notes*: Table entries are non-standardized random-effect GLS regression coefficients, clustered at the subject level. Standard errors are in parentheses. The reference category for the symbols manipulation conditions is the control no-symbols.",
          notes.append = FALSE)

@@

&ast;*p*<.1; &ast;&ast;*p*<.05; &ast;&ast;&ast;*p*<.01

@@

In Model 1.2, the effect of the symbols remains intact, and the effect of the weak policy condition is negative and significant (*Coef* = `r mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]). In Model 1.3, the interaction between weak policy and real symbols is positive and significant, whereas the interaction with the fake symbols is insignificant. The interpretation of this model suggests that the real symbols had a greater effect on trust when the policy plan was weak (*Coef* = `r mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`] and `r mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`] Compared with the no symbols and fake symbols, respectively). When added to the strong policy plan, the effect of the real symbols is weaker, and significant only compared with the no symbols condition (*Coef* = `r mod_1.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)`, *p*<.1). Congruently, the symbols decreased the differences in trust between the strong and weak policies by `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())/(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())*100)%>%round(0)`%, compared with the no symbols group (from `r mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`] to `r mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`]). The interaction between symbols and information significantly contributes to the model fit (Chi2 = `r anova(mod_1.3,mod_1.2)%>%tidy()%>%filter(term=="mod_1.3")%>%select(statistic)%>%max()%>%round(2)`, *p*<.05). To facilitate the interpretation of this interaction, **Figure 4** provides a descriptive demonstration of the effects of the real symbols on trust in policy, under weak and strong information quality.

@@

Altogether, the above between- and within-subject analyses largely support my first two hypothesis. They suggest that familiar symbolic elements in communications increase citizens' trust in policies which are presented, and that the effect is greater the less logically persuasive is the information in the communications. Symbols attenuate the negative effect of weak or logically unpersuasive information in communications.In other statistical models, shown in the supplementary appendix, I tested additional specifications, including ##TBA##. In all these robust analyses, the effect of the real symbols condition on trust in policy remains positive and significant, and the interaction bewteen the real symbols and the weak information remains positive and signifcant [##TBA##].

How can we explain the finding that the real symbols did not increase participants' trust in the strong policy (Model 1.3)? One tentative explanation could be that symbolic elements are more effective in shifting people from low to medium levels of trust, than shifting them from medium to high levels. Stated otherwise, symbols may be more effective in reducing citizens' distrust, as opposed to increasing their trust. If this explanation holds, then result will be a cielling effect. This explantion can be also linked, theoretically, to the notion of public sector negative bias, which is well established in the Public Administration. Since most citizens tend to have implicit negative attitudes towards the public sector, they can trust public organizations up until a certain degree.[^footnote_ceilling_effect]

[^footnote_ceilling_effect]: This explanation is partly supported by a descriptive comparison of the distributions of the different groups. The comparison of the quartiles in the different groups indicates that the increase in the average was mainly due to the increase in the first quartile (i.e. the lower 25% observations). In the weak policy condition, the first quartile increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==0])[2]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==0])[2])%>%max()%>%round(1)` points compared with the no symbols, whereas the third quartile (i.e. the higher 25% observations) increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==0])[4]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==0])[4])%>%max()%>%round(1)` points. In the strong policy condition, we can find similar pattern of decreasing marginal effect. The first quartile increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==1])[2]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==1])[2])%>%max()%>%round(1)` points, whereas the third quartile (i.e. the higher 25% observations) increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==1])[4]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==1])[4])%>%max()%>%round(1)` points.

@@

@@

```{r, echo=FALSE,warning=FALSE,message=FALSE, out.width=c('50%', '50%'),fig.show='hold',, fig.width=4,fig.height=4}
ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n), ELABORATION_time_log))+
  theme_classic()+
  #geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
                              fun.args = list(mult = 1.65),
                              width=0.1,size=0.5,color="gray70") + 
  stat_summary(fun.y = mean, geom = "point", size=1.5)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  scale_x_discrete(name="",labels=c("no symbols","fake symbols","real symbols"))+
  scale_y_continuous(name = "Reaction Time (log)")

@@

[^footnote_mediation_sensitivity]: I also conducted a robusteness test of "formal sensitivity" for the significant causal mediation models, as proposed by Imai et al. [-@imai_2010]. The sensitivity test examines the extent to which the causal mediation would change as a result of a correlation between the error terms of the mediator and the outcome variable (the "sequential ignorability" assumption). With regard to mediation by reaction time, the sensitivity factors range bewteen 0.2-0.3, which suggest that a correlation greater than these values would violate the sequential ignorability assumption. Imai et al. did not offer any rule of thumb with regard to these factors. Yet, these results are consistent with sensitivity factors reported in previous studies, and therefore the mediation can be considered fairly reliable. The sensitivity factors for the mediation by the memory score are 0.1, which entails that it is less reliable.

Altogether, the between- and within-subjects analyses are consistent with my third and fourth hypotheses. They demonstrate that subjects who saw the policy plans with the familiar symbols of EPM, were less likely to elaborate on the content of the EPM policies, which in turn attenuated the negative effect of the weak information, and increased their trust in the weak policies. The partial mediation by reaction time is strongly significant, whereas the mediation by the memory test is notsufficiently significant in all models. In some of these models, the familiar symbols also had a significant direct positive effect, which can be theoretically explained by the peripheral effect of symbols via trasfer of affect, in accordance with my first hypothesis.

>>>>

---
title: "The effect of symbols in government communications on citizens' trust, and its interaction with substantive information"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
urlcolor: blue
---

@@

Last edited at `r Sys.Date()`.[^footnote_scripts]

@@

Elements in communications can also function as factors that affect people's motivation and/or ability to think. For example, Background music in commercials can distract people and prevent them from scrutinizing their content [@park_1986]; The use of humor in television shows stimulates positive mood and reduces people's ability and motivation to scrutinize their substantive arguments [@nabi_2007; @young_2008]. Hence, elements can have more than one function, and can shape attitudes through several mechanisms. They can serve both as variables that affect the amount of thinking and as potential peripheral cues or arguments. Continuing with the above-mentioned examples, background music and humor can serve as simple cues that persuades people directly through the peripheral route, yet at the same time they can also as factors that reduce elaboration. Accordingly, they can affect people both directly (as peripheral cues) and indirectly (by decreasing elaboration and thus attenuating the effect of arguments and enhancing the effect of peripheral cues).[^footnote_multifunction_factors]

[^footnote_multifunction_factors]: @petty_1986 have suggested that we are more likely to find complex relations of multiple mechanisms in those situations where people are located in the middle of the elaboration continuum. When people are neither extremely motivated or unmotivated to think, they may rely on different elements in the communications to decide if the message is worth or needs considering.

@@

Analytically, the expectation that symbolic elements decrease message elaboration also entails that symbols may decrease citizens' trust in communications about highly positive performance and persuasive policy plans. Nevertheless, one can make the argument that neutralizing the effect of "positive" information with symbols may not actually yield significant differences, since citizens are more likely to "punish" bureaucracies for publishing negative information than to "praise" them for publishing positive information [@james_2011; @nielsen_2017; @james_2017b]. Congruently, the effect of positive information (versus a medium level) is likely to be relatively small, and hence neutralizing this effect may not make much difference. Given the focus of this study on the interaction between symbols and ill-conceived policies, the research design does not directly addresses this latter interaction.

@@

In addition to the above-mentioned hypotheses, the experiment is also designed to test an additional set of hypotheses regarding the interaction between symbols and the perceived personal relevance of communications. The selection of policy plans and the construction of the population sample were designed to account for variation in that variable. In this paper, I focus on those methodological and empirical parts of the study that are consequential for the examination of the interaction between symbols and information.[^footnote_relevance_paper]

[^footnote_relevance_paper]: The methodological aspects which are relevant for examining the additional hypotheses (i.e. the interaction between symbols with personal relevance) are described in a separate paper. An English translation of the full survey is available from the supplementary appendix.

@@

[^footnote_prior_studies]: In two prior online survey tests, which I conducted before the experiment, I collected further information about the affect and symbolic associations of these symbols. These results, which are reported in the supplementary appendix, substantiate my empirical assumption that these elements evoke positive emotions Israeli citizens, and that citizens automatically associate them with EPM and its policies. These results also indicate that the images of the celebrity comedians from EPM campaigns are perceived as humoristic, and accordingly, they are expected to position people in a positive mood.

@@

[^footnote_relevance_manipulation]: This section also included a manipulation of participants' perceived personal relevance. Participants in the treatment group were asked a series of questions about environmental government policy, whereas participants in the control group were asked instead equivalent questions about their field of occupation.

@@

[^footnote_symbols_manipulation_checks]: In two prior surveys, I empirically confirmed that the fake symbols are unfamiliar, do not have strong positive or negative affect, and are not strongly associated with environment policy and EPM. The results of these examinations are reported in the supplementary appendix.

@@

[^footnote_combined_analyses]: In the supplementary appendix I further replicate the analyses below, while adding the interactions between the manipulations and the policy plans. In these models, the interactions between the policy plans and the symbols are far from significance, with no major changes to the results.

@@

My main independent variables are the above mentioned manipulations for information and symbols, and the interaction between them. My outcome variable is citizens' trust in policy (or the perceived trustworthiness of the policy). This variable is measured in the survey for each policy plan using a composite index of six likert scale items (Cronbach’s alpha = `r alpha_waste_trust`), which were adopted and slightly modified from validated trust scales used by previous public administration studies on trust in government [@grimmelikhuijsen_2013; @grimmelikhuijsen_2014; @karens_2016; @porumbescu_2016; @grimmelikhuijsen_2017]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7(strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan. The mean of this trust index is `r SVIVA2_01_comb%>%summarise(mean(trust))%>%max()%>%round(2)`, the standard variation is `r SVIVA2_01_comb%>%summarise(sd(trust))%>%max()%>%round(2)` and the median is `r SVIVA2_01_comb%>%summarise(median(trust))%>%max()%>%round(2)`. Higher levels of trust are associated with higher trust in government, left-wing ideology and lower levels of education. In addition to measuring participants trust in each policy, per each subject, I also calculated the delta between her trust in the strong and weak policy plans which were presented to her (*Mean* = `r SVIVA2_01%>%summarise(mean(WITHIN_DELTA))%>%max()%>%round(2)`, *SD* = `r SVIVA2_01%>%summarise(sd(WITHIN_DELTA))%>%max()%>%round(2)`, *Median* = `r SVIVA2_01%>%summarise(median(WITHIN_DELTA))%>%max()%>%round(2)`). I used this Trust in policy delta to analyze the interaction between symbols and information within subjects.

@@

In addition to the above independent variables, the survey also included control items regarding respondents’ age, gender, education, income, home ownership, age of their children, political ideology, trust in government, interest in environmental issues and residence in the Haifa-Bay area. Summary statistics and correlation matrix for all variables are available in the supplementary appendix.

@@

I now turn to analyzing the experimental results. I begin by examining my first and second hypotheses, that regard the effect of symbols on trust and its interaction with information. Thereafter, I will examine hypotheses 3 and 4, that regard the mediating role on elaboration. For each section, I conducted both *between-subjects* and *within-subjects* analyses.

@@

**Table 1: Regression analyses – the effect of symbols on trust in policy**

@@

stargazer(mod_1.1,mod_1.2,mod_1.3,
          type="html",style = "apsr",
          dep.var.labels=c("Trust in policy (combined)"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Weak policy (0=strong policy)",
                             "Weak policy x Real symbols",
                             "Weak policy x Fake symbols"),
          initial.zero = FALSE,
          omit.stat = c("aic","bic"),
          notes = "*Notes*: Table entries are non-standardized random-effect GLS regression coefficients, clustered at the subject level. Standard errors are in parentheses. The reference category for the symbols manipulation conditions is the control no-symbols.",
          notes.append = FALSE)

@@

&ast;*p*<.1; &ast;&ast;*p*<.05; &ast;&ast;&ast;*p*<.01 (two-tailed).

@@

In Model 1.2, the effect of the symbols remains intact, and the effect of the weak policy condition is negative and significant (*Coef* = `r mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]). In Model 1.3, the interaction between weak policy and real symbols is positive and significant, whereas the interaction with the fake symbols is insignificant. The interpretation of this model suggests that the real symbols had a greater effect on trust when the policy plan was weak (*Coef* = `r mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`] and `r mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`] Compared with the no symbols and fake symbols, respectively). When added to the strong policy plan, the effect of the real symbols is weaker, and significant only compared with the no symbols condition (*Coef* = `r mod_1.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)`, *p*<.1, one-tailed). Congruently, the symbols decreased the differences in trust between the strong and weak policies by `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())/(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())*100)%>%round(0)`%, compared with the no symbols group (from `r mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`] to `r mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`]). The interaction between symbols and information significantly contributes to the model fit (Chi2 = `r anova(mod_1.3,mod_1.2)%>%tidy()%>%filter(term=="mod_1.3")%>%select(statistic)%>%max()%>%round(2)`, *p*<.05). To facilitate the interpretation of this interaction, **Figure 4** provides a descriptive demonstration of the effects of the real symbols on trust in policy, under weak and strong information quality.

@@

Altogether, the above between- and within-subject analyses largely support my first two hypothesis. They suggest that familiar symbolic elements in communications increase citizens' trust in policies which are presented, and that the effect is greater the less logically persuasive is the information in the communications. Symbols attenuate the negative effect of weak or logically unpersuasive information in communications. In other statistical models, shown in the supplementary appendix, I tested additional specifications, including analyzing the unfilterred sample, adding control variables, excluding subjects who did not recognize the real symbols and adding interactions with the policy plans. In all these robust analyses, the main effect of the real symbols remains positive and significant, and the interaction between the real symbols and the weak information remains positive and significant.

How can we explain the finding that the real symbols only had a marginally significant effect on participants' trust in the strong policy, compared with the no logo, and did not increase their trust compared with the fake logo (Model 1.3)? One tentative explanation could be that symbolic elements are more effective in shifting people from low to medium levels of trust, than shifting them from medium to high levels. Stated otherwise, symbols may be more effective in reducing citizens' distrust, as opposed to increasing their trust. If this explanation holds, then result will be a ceiling effect. This explanation can be also linked, theoretically, to the notion of public sector negative bias, which is well established in the Public Administration. Since most citizens tend to have implicit negative attitudes towards the public sector, they can trust public organizations up until a certain degree.[^footnote_ceiling_effect]

[^footnote_ceiling_effect]: This explanation is partly supported by a descriptive comparison of the distributions of the different groups. The comparison of the quartiles in the different groups indicates that the increase in the average was mainly due to the increase in the first quartile (i.e. the lower 25% observations). In the weak policy condition, the first quartile increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==0])[2]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==0])[2])%>%max()%>%round(1)` points compared with the no symbols, whereas the third quartile (i.e. the higher 25% observations) increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==0])[4]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==0])[4])%>%max()%>%round(1)` points. In the strong policy condition, we can find similar pattern of decreasing marginal effect. The first quartile increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==1])[2]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==1])[2])%>%max()%>%round(1)` points, whereas the third quartile (i.e. the higher 25% observations) increased by `r (quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==2&SVIVA2_01_comb$INFORMATION==1])[4]- quantile(SVIVA2_01_comb$trust[SVIVA2_01_comb$SYMBOL==0&SVIVA2_01_comb$INFORMATION==1])[4])%>%max()%>%round(1)` points.

@@

**Table 2: Regression analyses – the effect of symbols on elaboration**

@@

```{r, echo=FALSE,warning=FALSE,message=FALSE, out.width=c('50%', '50%'),fig.show='hold', fig.width=4,fig.height=4}
ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n), ELABORATION_time_log))+
  theme_classic()+
  #geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
                              fun.args = list(mult = 1.65),
                              width=0.1,size=0.5,color="gray70") + 
  stat_summary(fun.y = mean, geom = "point", size=1.5)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  scale_x_discrete(name="",labels=c("no symbols","fake symbols","real symbols"))+
  scale_y_continuous(name = "Reaction Time (log)")

@@

[^footnote_mediation_sensitivity]: I also conducted a robustness test of "formal sensitivity" for the significant causal mediation models, as proposed by Imai et al. [-@imai_2010]. The sensitivity test examines the extent to which the causal mediation would change as a result of a correlation between the error terms of the mediator and the outcome variable (the "sequential ignorability" assumption). With regard to mediation by reaction time, the sensitivity factors range between 0.2-0.3, which suggest that a correlation greater than these values would violate the sequential ignorability assumption. Imai et al. did not offer any rule of thumb with regard to these factors. Yet, these results are consistent with sensitivity factors reported in previous studies, and therefore the mediation can be considered fairly reliable. The sensitivity factors for the mediation by the memory score are 0.1, which entails that it is less reliable.

Altogether, the between- and within-subjects analyses are consistent with my third and fourth hypotheses. They demonstrate that subjects who saw the policy plans with the familiar symbols of EPM, were less likely to elaborate on the content of the EPM policies, which in turn attenuated the negative effect of the weak information, and increased their trust in the weak policies. The partial mediation by reaction time is strongly significant, whereas the mediation by the memory test is not sufficiently significant in all models. In some of these models, the familiar symbols also had a significant direct positive effect, which can be theoretically explained by the peripheral effect of symbols via transfer of affect, in accordance with my first hypothesis.
